---
title: "Midterm2"
author: "Elliot Metviner"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
sim <- function(i) {
  length <- 1
  v <- c()
    while(i != 1) {
      v[length] = i
      length <- length + 1
      if(i%%2 == 0){
        i <- i/2
      } else {
          i <- i*3 + 1 
          }
    }
  length
}
```

```{r}
out <-c()
for (i in 1:10000) {
  out[i] <- sim(i)
}
out

```


```{r}
hist(out)
```
```{r}
summary(out)
```

**The distribution of the graph is roughly normal but is skewed to the right. this matches the summary statstic that has a minimum of 1 and a maximum of 262.0, with a median of 74.**


```{r}

```


**I was unable to tabulate but I bet that even numbers occured more frequently than odd numbers (other than one). I know this because when we encounter an odd number we make it even for the next value, and when we divide a number there is a chance it can be odd but also can be even.  **



